---
title: "Replication of results"
subtitle: '"Human-AI Interactions in Public Sector Decision-Making"'
author: "Saar Alon-Barkat"
date: "January 2022"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 3
---



```{r set-global-options, echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = TRUE, 
                      message=FALSE,
                      warning = FALSE,
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      dependson = NULL,
                      engine = "R", #  Chunks will always have R code, unless noted
                      error = TRUE)
```

```{r}
library(tidyverse) # version 1.3.1
library(gtsummary) # version 1.4.1
library(sjPlot) # version 2.8.7
```



# Data preparations


Import datasets (unfiltered) for studies 1,2,3. 
```{r}
study1_unfiltered <-
  read_csv("data_for_publication/study1_unfiltered.csv")

study2_unfiltered <-
  read_csv("data_for_publication/study2_unfiltered.csv")

study3 <- read_csv("data_for_publication/study3.csv")
```



Filtering of Participants who failed the attention check or completed the questionnaire in less than three minutes (studies 1 & 2). 
```{r}
study1 <- study1_unfiltered %>%
  filter(IMC == 1, timer.minutes >= 3)

study2 <- study2_unfiltered %>%
  filter(IMC == 1, timer.minutes >= 3)

```


Creating pooled dataset for studies 1 & 2

```{r}

pooled <- study1 %>%
  full_join(study2) 
```


Participants not of Dutch descent (excluded from the analysis of selective adherence hypotheses - H2 & H3).
```{r}
study2.selective <- study2 %>%
  filter(excluded.selective == 0)

study3.selective <- study3 %>%
  filter(excluded.selective == 0)

```


<br>


# Datasets and variables

**study 1**
```{r}
study1 %>% glimpse()
```

<br>


**Study 2**
```{r}
study2 %>% glimpse()
```
<br>

**Study 3**
```{r}
study3 %>% glimpse()
```

<br>

# Main analyses


<br>


## Study 1

### Table 1

```{r}
mod1 <- glm(follow.advice ~ advice.algorithm,
            family = "binomial",
            data = study1)

sjPlot::tab_model(
  mod1,
  show.stat = T,
  show.loglik = T,
  show.r2 = F,
  collapse.ci = T,
  emph.p = F,
  string.est = "OR (95% CI)",
  string.stat = "z",
  string.p = "p-value",
  digits.p = 3,
  order.terms = c(2, 1)
)


```


<br>


### Table 2

```{r}

study1 %>%
  select(decision.comb,
         advice) %>%
  gtsummary::tbl_summary(by = advice,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")

```


<br>

## Study 2

*Automation bias* 

### Table 3

```{r}
mod1 <- glm(follow.advice ~ advice.algorithm,
            family = "binomial",
            data = study2)

mod2 <- glm(follow.advice ~ advice.algorithm + study,
            family = "binomial",
            data = pooled)



sjPlot::tab_model(
  mod1,
  mod2,
  show.stat = T,
  show.loglik = T,
  show.r2 = F,
  collapse.ci = T,
  emph.p = F,
  string.est = "OR (95% CI)",
  string.stat = "z",
  string.p = "p-value",
  digits.p = 3,
  order.terms = c(2:3, 1)
)

```

<br>


### Table 4


**Study 2**

```{r}

study2 %>%
  select(decision.comb,
         advice) %>%
  gtsummary::tbl_summary(by = advice,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")
 
```

**Studies 1 & 2 (pooled)**

```{r}
pooled %>% 
  select(decision.comb,
         advice) %>% 
  gtsummary::tbl_summary(by = advice,
              digits = list(all_categorical() ~ c(0, 1)))%>%
  gtsummary::modify_header(label ~ "**Teacher selected**
                (non-renewal of employment)")
```

<br>


*Selective adherence*

### Table 5

```{r}
mod1 <- glm(
  follow.advice ~ advice.algorithm+moroccan.teacher.cond.bi.lab,
  family = "binomial",
  data = study2.selective
)

mod2 <- glm(
  follow.advice ~ advice.algorithm*moroccan.teacher.cond.bi.lab,
  family = "binomial",
  data = study2.selective
)



sjPlot::tab_model(mod1,mod2, 
          show.stat = T,
          show.loglik = T,
          show.r2 = F,
          collapse.ci = T,
          emph.p = F,
          string.est = "OR (95% CI)",
          string.stat = "z",
          string.p = "p-value",
          digits.p = 3,
          order.terms = c(2:4,1),
          show.aic = T)

```


<br>


### Table 6

**All**
```{r}

study2.selective %>% 
  select(moroccan.teacher.cond.bi.lab,
         decision.comb) %>% 
  gtsummary::tbl_summary(by = moroccan.teacher.cond.bi.lab,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")


```

**By advice**

```{r}


study2.selective %>% 
  select(decision.comb,
         advice,
         moroccan.teacher.cond.bi.lab) %>%   
  gtsummary::tbl_strata(
    strata = advice,
    .tbl_fun =
      ~ .x %>%
        gtsummary::tbl_summary(by = moroccan.teacher.cond.bi.lab,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")
  )
```


<br>


## Study 3

*Automation bias* 

### Table 7

```{r}
mod1 <- glm(follow.advice ~ advice.algorithm,
            family = "binomial",
            data = study3)






sjPlot::tab_model(
  mod1,
  show.stat = T,
  show.loglik = T,
  show.r2 = F,
  collapse.ci = T,
  emph.p = F,
  string.est = "OR (95% CI)",
  string.stat = "z",
  string.p = "p-value",
  digits.p = 3,
  order.terms = c(2, 1)
)

```


<br>

### Table 8



```{r}

study3 %>%
  select(decision.comb,
         advice) %>%
  gtsummary::tbl_summary(by = advice,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")
 
```

<br>


*Selective adherence*

### Table 9

```{r}
mod1 <- glm(
  follow.advice ~ advice.algorithm + moroccan.teacher.cond.bi.lab,
  family = "binomial",
  data = study3.selective
)

mod2 <- glm(
  follow.advice ~ advice.algorithm * moroccan.teacher.cond.bi.lab,
  family = "binomial",
  data = study3.selective
)




sjPlot::tab_model(
  mod1,
  mod2,
  show.stat = T,
  show.loglik = T,
  show.r2 = F,
  collapse.ci = T,
  emph.p = F,
  string.est = "OR (95% CI)",
  string.stat = "z",
  string.p = "p-value",
  digits.p = 3,
  order.terms = c(2:4, 1),
  show.aic = T
)

```


<br>


### Table 10

**All**
```{r}

study3.selective %>%
  select(moroccan.teacher.cond.bi.lab,
         decision.comb) %>%
  gtsummary::tbl_summary(by = moroccan.teacher.cond.bi.lab,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
  gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")


```


**By advice**

```{r}

study3.selective %>%
  select(decision.comb,
         advice,
         moroccan.teacher.cond.bi.lab) %>%
  gtsummary::tbl_strata(
    strata = advice,
    .tbl_fun =
      ~ .x %>%
      gtsummary::tbl_summary(by = moroccan.teacher.cond.bi.lab,
                         digits = list(all_categorical() ~ c(0, 1)),
                         label = decision.comb ~ "") %>%
      gtsummary::modify_header(label ~ "**Teacher selected** (non-renewal of employment)")
  )
```


<br>


## Sample characteristics

```{r}
t1 <- study1 %>%
  full_join(study2) %>%
  full_join(study3) %>%
  mutate(
    age_cat = case_when(
      age <= 25 ~ "18-25",
      age > 25 & age <= 35 ~ "26-35",
      age > 35 & age <= 45 ~ "36-45",
      age > 45 & age <= 55 ~ "46-55",
      age > 55 & age <= 65 ~ "56-65",
      age > 65 ~ "65+"
    )
  )


t1 %>%
  select(study,
         female,
         age,
         age_cat,
         high.education) %>%
  gtsummary::tbl_summary(
    by = study,
    statistic = list(all_numeric() ~ "{mean} ({sd})",
                     all_categorical() ~ "{p}%"),
    digits = list(all_categorical() ~ c(1),
                  all_continuous() ~ c(1, 1))
  )
```

